\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-04-17 doc/paper.tex
document minimum required g++ version (4.4.7)
explain allowed constraint forms

\section{Download and compilation}
Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
To use FlexibleSUSY please make sure you have the following installed.  
\begin{itemize}
\item Mathematica, version 7 or higher
\item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
\item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or
  higher)
\item Fortran compiler (gfortran, ifort etc.)
\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
\item Boost library, version 1.36.0 or higher
  \url{http://www.boost.org}
\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
\item lapack (needed for the Lattice algorithm only)
  \url{http://www.netlib.org/lapack/}
\end{itemize}
%
Optional:
%
\begin{itemize}
\item Looptools, version 2.8 or higher
  \url{http://www.feynarts.de/looptools/}
\end{itemize}

@@

Furthermore, the user has to specify the form of three model
constraints: low-scale, susy-scale and high-scale constraints.  In \fs
they are named as \code{LowScale}, \code{SUSYScale} and
\code{HighScale}.  For each constraint there is a scale definition
(named after the constraint), an initial guess for the scale
(concatenation of the constraint name and \code{FirstGuess}) and a
list of settings to be applied at the constraint (concatenation of the
constraint name and \code{Input}).
%
\begin{itemize}
\item \emph{High-scale constraint:} The high-scale constraint is
  usuall the GUT-scale constraint, applied at the scale where the
  gauge couplings $g_1$ and $g_2$ unify.  The high-scale can be
  defined by an equation of the form \code{g1 == g2} or by fixed scale
  value.
\item \emph{Susy-scale constraint:} The susy-scale constraint is the
  constraint where \fs will impose the EWSB conditions and calculate
  the spectrum.  It is usually defined by
  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.
\item \emph{Low-scale constraint:} The low-scale constraint is the
  constraint where the gauge couplings $g_1$, $g_2$, $g_3$ and the
  Yukawa couplings $y_f$ will be calculated from given Standard Model
  input $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $m_f^\pole$,
  $M_Z^\pole$, $M_W^\pole$, see
  Sec.~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
  This should be done at the $Z$ mass scale $M_Z$.  Note that the
  gauge and Yukawa couplings are calculated automatically and
  therefore do not need to be specified in the \code{LowScaleInput}
  variable.
\end{itemize}
%
Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
high-scale definition takes the simple form \code{g1 == g2}.  Another
consequence is that in the calculation of the VEVs $v_u$ and $v_d$
from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
to be taken into account, see above.

>>>>

\section{Download and compilation}
Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
To use FlexibleSUSY please make sure you have the following installed.  
\begin{itemize}
\item Mathematica, version 7 or higher
\item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
\item C++11 compatible compiler (g++ 4.4.7 or higher, clang++ 3.1 or
  higher)
\item Fortran compiler (gfortran, ifort etc.)
\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
\item Boost library, version 1.36.0 or higher
  \url{http://www.boost.org}
\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
\item lapack (needed for the Lattice algorithm only)
  \url{http://www.netlib.org/lapack/}
\end{itemize}
%
Optional:
%
\begin{itemize}
\item Looptools, version 2.8 or higher
  \url{http://www.feynarts.de/looptools/}
\end{itemize}

@@

Furthermore, the user has to specify the form of three model
constraints: low-scale, susy-scale and high-scale constraints.  In \fs
they are named as \code{LowScale}, \code{SUSYScale} and
\code{HighScale}.  For each constraint there is (i) a scale definition
(named after the constraint), (ii) an initial guess for the scale
(concatenation of the constraint name and \code{FirstGuess}) and (iii)
a list of parameter settings to be applied at the scale (concatenation
of the constraint name and \code{Input}).
%
\begin{itemize}
\item \emph{High-scale constraint:} The high-scale constraint is
  usuall the GUT-scale constraint, applied at the scale where the
  gauge couplings $g_1$ and $g_2$ unify.  The high-scale can be
  defined by an equation of the form \code{g1 == g2} or by fixed scale
  value.
\item \emph{Susy-scale constraint:} The susy-scale constraint is the
  constraint where \fs will impose the EWSB conditions and calculate
  the spectrum.  It is usually defined by
  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.
\item \emph{Low-scale constraint:} The low-scale constraint is the
  constraint where the gauge couplings $g_1$, $g_2$, $g_3$ and the
  Yukawa couplings $y_f$ will be calculated from given Standard Model
  input $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $m_f^\pole$,
  $M_Z^\pole$, $M_W^\pole$, see
  Sec.~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
  This should be done at the $Z$ mass scale $M_Z$.  Note that the
  gauge and Yukawa couplings are calculated automatically and
  therefore do not need to be specified in the \code{LowScaleInput}
  variable.
\end{itemize}
%
Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
high-scale definition takes the simple form \code{g1 == g2}.  Another
consequence is that in the calculation of the VEVs $v_u$ and $v_d$
from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
to be taken into account, see above.

The list of parameter settings can contain as elements any of the
following:
%
\begin{itemize}
\item Two-component lists of the form \code{\{parameter, value\}},
  which indicates that the \code{parameter} is set to \code{value} at
  the defined scale.  If the \code{value} should be read from the SLHA
  input file, it must be written as \code{LHInput[value]}.  Example:
  %
  \begin{lstlisting}
SUSYScaleInput = {
   {mHd2, m0^2},
   {mHu2, LHInput[mHu2]}
};
  \end{lstlisting}
  %
  In this example the parameter \code{mHd2} is set to the value of
  \code{m0^2}, and \code{mHu2} is set to the value given in the SLHA
  input file in block \code{MSOFTIN}, entry 22.  The SLHA block names
  and keys for the MSSM and NMSSM are defined in SARAH's
  \code{parameters.m} file, see the SARAH manual or
  \cite{Staub:2010jh}.

\item The function \code{FSMinimize[parameters, function]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{function} is a function of these parameters.
  \code{FSMinimize[parameters, function]} will numericall vary the
  \code{parameters} until the \code{function} is minimized.  Example:
  %
  \begin{lstlisting}
FSMinimize[{vd,vu},
           (SM[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
           (SM[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  function
  %
  \begin{align}
    \chi^2(v_d,v_u) =
    \frac{(\texttt{SM[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
    \frac{(\texttt{SM[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
  \end{align}
  %
  is minimal.  The constants \code{SM[MZ]}, \code{SM[MH]},
  $\sigma_{m_Z}$ and $\sigma_{m_h}$ are defined in
  \code{src/ew_input.hpp} to be
  %
  \begin{align}
    \texttt{SM[MZ]} &= 91.1876, &
    \texttt{SM[MH]} &= 125.9, \\
    \sigma_{m_Z} &= 0.0021, &
    \sigma_{m_h} &= 0.4 .
  \end{align}

\item The function \code{FSFindRoot[parameters, functions]} can be
  given, where \code{parameters} is a list of model parameters and
  \code{functions} is a list of functions of these parameters.
  \code{FSFindRoot[parameters, functions]} will numericall vary the
  \code{parameters} until the \code{functions} are zero.  Example:
  %
  \begin{lstlisting}
FSFindRoot[{vd,vu},
           {SM[MZ] - Pole[M[VZ]], SM[MH] - Pole[M[hh[1]]]}]
  \end{lstlisting}
  %
  Here, the parameters \code{vu} and \code{vd} are varied until the
  vector-like function
  %
  \begin{align}
    f(v_d,v_u) =
    \begin{pmatrix}
      \texttt{SM[MZ]} - m_Z^\pole \\
      \texttt{SM[MH]} - m_{h_1}^\pole
    \end{pmatrix}
  \end{align}
  %
  is zero.
\end{itemize}
